### 2021-12-14 thesis/tex/sterileTheory.tex
update

%\input{commands}
%\begin{document}
\section{The Sterile State}\label{sec:anomalies}
In 1996, the LSND experiment reported an excess of $\ane$ events from an $\anm$ beam~\cite{lsnd}.  
Nine years later, MiniBooNE not only reproduced the $\ane$ anomaly but observed an excess in the
$\ne$ events too. Together with the so-called reactor and gallium anomalies, these reports suggested 
that the three massive neutrino framework could be amended.
By introducing a fourth heavy mass state $\nu_4$, both appearance and disappearance anomalies could be minimally accommodated.
However, we know from the decay width of the $Z$ boson that it only can interact with three flavor species\footnote{
    Direct measurements of the invisible $Z$ decay yields $N=2.92 \pm 0.05$~\cite{pdg}.
}, 
so this fourth mass state cannot be interacting weakly. In other words, it needs to transform as a
singlet under the gauge symmetry group $SU(2)_L \times U(1)_{Y}$.
We now distinguish between the three original neutrino flavors ($e$, $\mu$, and $\tau$) and the new fourth 
flavor by calling the former \emph{active} neutrinos
and the latter \emph{sterile} due to its non-interacting behavior.

@@

Keeping in mind that the event count from Eq.~\ref{eq:ICevents} scales as $E^{-1.7}$,
signals at higher energies will be more difficult to detect. Since higher values of $\dm[41]$ shifts the $\Pamam$ resonance peak to higher 
energies, high $\dm[41]$ values will be harder to rule out because the event count is already lower, as compared to energies 
at single-digit \si{\TeV} values. There is then a range of parameters, in which our detector will 
be able to distinguish between the two hypotheses. Using the two-neutrino approximation\footnote{In this approximation, we only consider mixing between $\nm$ and $\ns$, essentially a `1+1' picture.}, 
we can rewrite the expression for MSW resonance from Eq.~\ref{eq:MSW} as 
\begin{align}
    E^* = \frac{\dm[41] \cos{2\theta_{24}}}{2\sqrt{2} G_F N_e}\,,
\end{align}
which for $\dm[41] = \SI{1}{\TeV^2}$ and $\theta_{24} = 0.04$, gives us a resonance at $E^* = \SI{2.58}{\TeV}$, matching the $3+1$ simulation shown in
Fig.~\ref{fig:sterile_resonance}. From the two-neutrino picture, 
we see that the resonance peak in the $\anm$ channel scales linearly with $\dm[41]$, while the mixing angle dependence 
shows up as $\cos{2\theta_{24}}$, which is between 0.987 and 0.999 for the $\theta_[24]$ values considered in Fig.~\ref{fig:resonance_shift}.
Since our $\theta_{24}$ values considered are so low, the location of the resonance peak is essentially only dependent on the value of $\dm[41]$.
Thus, we expect a more stringent constraint of low values of $\dm[41]$, regardless of the value of $\theta_{24}$.

>>>>

\section{The Sterile State}\label{sec:anomalies}
In 1996, the LSND experiment reported an excess of $\ane$ events from an $\anm$ beam~\cite{lsnd}.  
Nine years later, MiniBooNE not only reproduced the $\ane$ anomaly but observed an excess in the
$\ne$ events too. Together with the so-called reactor and gallium anomalies, these reports suggested 
that the three massive neutrino framework could be amended.
By introducing a fourth heavy mass state $\nu_4$, both appearance and disappearance anomalies could be minimally accommodated.
However, we know from the decay width of the $Z$ boson that it only can interact with three flavor species\footnote{
    Direct measurements of the invisible $Z$ decay yields $N=2.92 \pm 0.05$~\cite{pdg}.
}, 
so this fourth mass state cannot be interacting weakly. In other words, it needs to transform as a
singlet under the gauge symmetry group $SU(2)_L \times U(1)_{Y}$.
We now distinguish between the three original neutrino flavors ($e$, $\mu$, and $\tau$) and the new fourth 
flavor by calling the former \emph{active} neutrinos
and the latter \emph{sterile} due to its non-interacting behavior.

@@

Keeping in mind that the event count from Eq.~\ref{eq:ICevents} scales as $E^{-1.7}$,
signals at higher energies will be more difficult to detect. Since higher values of $\dm[41]$ shift the $\Pamam$ resonance peak to higher 
energies, high $\dm[41]$ values will be harder to rule out because the event count is already lower, as compared to energies 
at single-digit \si{\TeV} values. There is then a range of parameters with which our detector will 
be able to distinguish between the two hypotheses. Using the two-neutrino approximation\footnote{In this approximation, we only consider mixing between $\nm$ and $\ns$, essentially a `1+1' picture.}, 
we can rewrite the expression for MSW resonance from Eq.~\ref{eq:MSW} as 
\begin{align}
    E^* = \frac{\dm[41] \cos{2\theta_{24}}}{2\sqrt{2} G_F N_e}\,,
\end{align}
which for $\dm[41] = \SI{1}{\TeV^2}$ and $\theta_{24} = 0.04$, gives us a resonance at $E^* = \SI{2.58}{\TeV}$, matching the $3+1$ simulation shown in
Fig.~\ref{fig:sterile_resonance}. From the two-neutrino picture, 
we see that the resonance peak in the $\anm$ channel scales linearly with $\dm[41]$, while the mixing angle dependence 
shows up as $\cos{2\theta_{24}}$, which is between 0.987 and 0.999 for the $\theta_[24]$ values considered in Fig.~\ref{fig:resonance_shift}.
Since our $\theta_{24}$ values considered are so low, the location of the resonance peak is essentially only dependent on the value of $\dm[41]$.
Thus, we expect a more stringent constraint of low values of $\dm[41]$, regardless of the value of $\theta_{24}$.
